---
title: "Miniproject"
author: "Poorvi Ashok"
date: "2023-02-13"
output: word_document
---


```{r libraries, results='hide'}
require(dplyr)
require(tidyr)
require(purrr)
require(ggplot2)
require(usdm)
require(knitr)
require(sjPlot)
require(ggeffects)
require(cowplot)
```
The dataset containing selected Environmental, Social and Governance (ESG) variables (mentioned below) for countries was downloaded from the World Bank Sovereign ESG Data Portal and input. Additionally, datasets containing other socio-political variables (mentioned below) for countries were also downloaded from the World Bank DataBank. Climate Change Performance Index (CCPI) data for countries was obtained and input from the CCPI website. The CCPI data was obtained in a pdf format and then converted into a csv format.

World Bank DataBank: https://databank.worldbank.org/source/environment-social-and-governance?preview=on. 
CCPI Website: https://ccpi.org/download/climate-change-performance-index-2023/

```{r input datasets, results='hide'}
#setting working directory
setwd("/Users/tp/Desktop/Miniproject")

#input ESG data
data_t <- read.csv('ESG.csv')
summary(data_t)
str(data_t)

#input CCPI data
CCPI <- read.csv('CCPI New.csv')
summary(CCPI)
str(CCPI)

#input national income per capita
NI <- read.csv('National Income Per Capita.csv')
summary(NI)
str(NI)

#input GDP per capita 
GC <- read.csv('GDP Per Capita.csv')
summary(GC)
str(GC)

#input population growth dataset
PG <- read.csv('Population Growth.csv')
summary(PG)
str(PG)

#input sex ratio dataset
SR <- read.csv('Sex Ratio.csv')
summary(SR)
str(SR)
```
Data wrangling was performed on the ESG dataset to contain the country name and corresponding values for the chosen ESG variables.

```{r data cleaning - ESG Dataset, results='hide'}
#ESG Data
#selecting only country name, ESG variable name and 2019 data from ESG dataset
data <- subset(data_t, select=-c(Country.Code, Series.Code, 
                                 X2013..YR2013.:X2018..YR2018., 
                                 X2020..YR2020.:X2022..YR2022.))
summary(data)
head(data)

#Pivoting the dataframe 
dataf <- pivot_wider(data, names_from = 'Series.Name', 
                        values_from = 'X2019..YR2019.')
summary(dataf)
head(dataf)

#renaming the colnames
names1 <- c('Country', 'Welfare_Aggregate_Growth', 'GDP_Growth', 
           'Digital_Use', 'Labour_Participation_Rate', 'Life_Expectancy', 
           'Literacy_Rate', 'Ageing_Population', 'Population_Density', 
           'Poverty_Ratio', 'Labour_Gender_Ratio', 'School_Enrollment', 
           'School_Enrollment_Gender_Ratio', 'Unemployment')
colnames(dataf) <- names1
head(dataf)

#Removing variables with low number of observations
dataf <- subset(dataf, select=-c(Welfare_Aggregate_Growth, Literacy_Rate, Poverty_Ratio))
summary(dataf)
head(dataf)
```
The ESG variables 'Welfare Aggregate Growth', 'Literacy Rate' and 'Poverty Ratio' were found to contain a low number of observations and were hence removed from the analysis.

Data wrangling was performed on the National Income dataset to contain the country name and the corresponding national income values.

```{r data manipulation - national income, results='hide'}
#Selecting onyl 2019 values
NI <- subset(NI, select=c('Data.Source', 'X.61'))
head(NI)

#Removing top rows (are either empty or not relevant)
NI <- NI[-(1:4),]
head(NI)

#renaming the columns
names3 <- c('Country', 'National_Income')
colnames(NI) <- names3
head(NI)
```
Data wrangling was performed on the GDP per Capita dataset to contain the country name and the corresponding GDP per Capita values.

```{r data manipulation - GDP per capita, results='hide'}
#Selecting only 2019 values
GC <- subset(GC, select=c('Data.Source', 'X.61'))
head(GC)

#Removing top rows (are either empty or not relevant)
GC <- GC[-(1:4),]
head(GC)

#renaming the columns
names3 <- c('Country', 'GDP_Capita')
colnames(GC) <- names3
head(GC)
```
Data wrangling was performed on the population growth dataset to contain the country name and the corresponding population growth values.

```{r data manipulation - population growth, results='hide'}
#Selecting only 2019 values
PG <- subset(PG, select=c('Country.Name', 'X2019'))
summary(PG)
head(PG)

#Changing column names 
names4 <- c('Country', 'Population_Growth')
colnames(PG) <- names4
head(PG)
```
Data wrangling was performed on the Sex Ratio dataset to contain the country name and the corresponding sex ratio values.

```{r data manipulation - sex ratio, results='hide'}
#Selecting only 2019 values
SR <- subset(SR, select=c('Country.Name', 'X2019..YR2019.'))
head(SR)

#Changing column names
names6 <- c('Country', 'Sex_Ratio')
colnames(SR) <- names6
head(SR)
```
Data wrangling was performed on the CCPI dataset to contain the country name and the corresponding CCPI values. 'Rank_Actual' was chosen instead of 'Rank' since it was consistent and started from 1. The choice of starting 'Rank' from 4 was made by the CCPI. This was done to reflect the lack of climate change mitigation performance by all countries, deeming none of their CCPI scores high enough to occupy the first 3 ranks.

```{r data manipulation CCPI dataset, results='hide'}
#Selecting only country,  actual rank & score
#CCPI <- subset(CCPI, select=c('X.1', 'X.2', 'X.3'))
CCPI <- subset(CCPI, select=c('Actual.Rank', 'Country', 'Score..'))
summary(CCPI)
head(CCPI)

#Removing the first 2 rows (Redundant rows)
#CCPI <- CCPI[-c(1:2),]
CCPI <- CCPI[-c(1:3),]
head(CCPI)

#Renaming the column names
names5 <- c('CCPI_Rank', 'Country', 'CCPI_Score')
colnames(CCPI) <- names5
head(CCPI)
```
Post data wrangling the ESG, CCPI, National Income, GDP per Capita, Population Growth and Sex Ratio datasets were merged to form the final dataset. The values were converted to numeric type variable and the data was cleaned to remove all NA values.

```{r final dataset, results='hide'}
#Merging the ESG and CCPI datasets
Total1<- merge(dataf, CCPI, by.x='Country', all.x=T)
#Merging Population Growth and Sex Ratio datsets
Total2 <- merge(SR, PG, by.x='Country', all.x=T)
#Merging National Income and GDP per Capita Datsets
Total3 <- merge(NI, GC, by.x='Country', all.x=T)
#Merging all datsets to create the final datset
Final <- merge(Total1, Total2, by.x='Country', all.x=T)
Final <- merge(Final, Total3, by.x='Country', all.x=T)
summary(Final)
head(Final)
str(Final)

#Converting all values in the dataframe to numeric
Final[,2:17] <- as.data.frame(sapply(Final[,2:17], as.numeric))
str(Final)

#omitting all NAs in the dataframe 
Final <- na.omit(Final)
```
The final dataset contained the values of 14 socio-economic and demographic variables, and CCPI rank and score for 48 countries. 

A linear model was chosen to study the effect of socio-economic and demographic variables on the CCPI scores of the countries. All the variables were checked for outliers.

```{r Checking linear model conditions - Plotting outliers, echo=TRUE}
par(mfrow=c(2,2))

#Plotting outliers 
boxplot(Final$CCPI_Score)
title(xlab='CCPI Score')
#No outliers
#2 outliers

boxplot(Final$School_Enrollment)
title(xlab='School Enrollment Ratio')
#Few outliers (logical)

boxplot(Final$Digital_Use)
title(xlab='Digital Literacy')
#2 outliers (logical)

boxplot(Final$School_Enrollment_Gender_Ratio)
title(xlab='School Enrollment Gender Ratio')
#1 outlier (logical)

boxplot(Final$Labour_Participation_Rate)
title(xlab = 'Labour Force Participation Rate')
#2 outliers (logical)

boxplot(Final$Unemployment)
title(xlab = 'Unemployment')
#4 outliers (logical)

boxplot(Final$Labour_Gender_Ratio)
title(xlab='Gender Ratio of Labour Participation')
#Few outliers (logical)

boxplot(Final$GDP_Growth)
title(xlab='GDP Growth')
#No outliers

boxplot(Final$National_Income)
title(xlab='National Income Per Capita')
#No outliers

boxplot(Final$GDP_Capita)
title(xlab='GDP per Capita')
#1 outlier (logical)

boxplot(Final$Life_Expectancy)
title(xlab='Life Expectancy')
#1 outliers (logical)

boxplot(Final$Ageing_Population)
title(xlab='Ageing Population')
#No outliers

boxplot(Final$Sex_Ratio)
title(xlab='Sex Ratio')
#3 outliers (logical)

boxplot(Final$Population_Growth)
title(xlab='Population Growth')
#No outliers

boxplot(Final$Population_Density)
title(xlab='Population Density')
#Few outliers (logical)
```
Outliers were found for the variables 'School_Enrollment', 'Digital_Use', 'School_Enrollment_Gender_Ratio', 'Labour_Participation_Rate', 'Unemployment', 'Labour_Gender_Ratio', 'GDP_per_Capita', 'Life_Expectancy', 'Sex_Ratio', and 'Population_Density'. The outliers were checked for abnormalities. All were found to be logical and hence were retained.

The homogeneity of the data was checked for all variables.

```{r checking linear model condition - homogeneity, echo=TRUE}
while (!is.null(dev.list()))  dev.off()

#Checking Homogeneity of Variances 
var(Final$CCPI_Score)
var(Final$Digital_Use, na.rm=T)
var(Final$School_Enrollment, na.rm=T)
var(Final$School_Enrollment_Gender_Ratio, na.rm=T)
var(Final$Labour_Participation_Rate, na.rm=T)
var(Final$Unemployment, na.rm=T)
var(Final$Labour_Gender_Ratio, na.rm=T)
var(Final$GDP_Growth, na.rm=T)
var(Final$GDP_Capita, na.rm=T)
var(Final$National_Income, na.rm=T)
var(Final$Life_Expectancy, na.rm=T)
var(Final$Ageing_Population, na.rm=T)
var(Final$Sex_Ratio, na.rm=T)
var(Final$Population_Growth, na.rm=T)
var(Final$Population_Density, na.rm=T)
#Doesn't satisfy conditions of homogeneity 
#Needs to be scaled 
```
The conditions for homogeneity were not satisfied by the variables. The variables would need to be scaled.

The conditions of normality were checked for the variables. 

```{r checking linear model conditions - normal distribution, echo=TRUE}
par(mfrow=c(2,2))

#Checking for normal distribution

hist(Final$CCPI_Score)
#normal 

hist(Final$School_Enrollment)
#normal (barring outliers)

hist(Final$Digital_Use)
#left-skewed data 

hist(Final$School_Enrollment_Gender_Ratio)
#normal

hist(Final$Labour_Participation_Rate)
#left-skewed

hist(Final$Unemployment)
#right-skewed

hist(Final$Labour_Gender_Ratio)
#left skewed

hist(Final$GDP_Capita)
#right skewed

hist(Final$GDP_Growth)
#normal 

hist(Final$National_Income)
#right-skewed

hist(Final$Life_Expectancy)
#left skewed

hist(Final$Ageing_Population)
#normal 

hist(Final$Sex_Ratio)
#normal (barring outliers) 

hist(Final$Population_Growth)
#normal 

hist(Final$Population_Density)
#right skewed
```
Variables 'Digital_Use', 'Labour_Participation_Rate', 'Labour_Gender_Ratio' and 'Life_Expectancy' were found to be left skewed while variables 'Unemployment', 'GDP_Capita', 'National_Income' and 'Population_Density were found to be right-skewed. The rest of the variables were found to contain a roughly normal distribution. The variables with skewed data should be kept in mind while checking the residual plots of the linear model. 

The variables were checked for data points containing 0s. 

```{r linear model conditions - zeroes, echo=TRUE}
colSums(Final==0)
```
There are no 0s in the dataframe. 

The collinearity among the explanatory variables was checked using the functions corr and variance inflation factor (VIF). 

```{r checking linear model conditions - collinearity/VIF, echo=TRUE}
#Checking for collinearity
pairs(Final[-c(1,12,13)])
cor(Final[-c(1,12,13)], y=NULL, use='complete.obs')

#Calculating VIF
vif(Final[-c(1,12,13)])
#Dropping National_Income
vif(Final[-c(1,12:13,16)])
#Dropping Ageing Population
vif(Final[-c(1,6,12:13,16)])
#Dropping Labour Gender Ratio
vif(Final[-c(1,6,8,12:13,16)])
#All good
```
Choosing a corr > 0.75 to be strong, the variable pairs 'Labour_Participation_Rate' and 'Labour_Gender_Ratio'; and 'GDP_Capita' and 'National_Income' were found to have a strong correlation. With the consideration of VIF > 5 indicating high correlation, the variables 'National_Income', 'Ageing_Population' and 'Labour_Gender_Ratio' were found to have a high correlation. Therefore, the variables 'National_Income', 'Ageing_Population' and 'Labour_Gender_Ratio' were omitted from the linear model.

A simple linear regression was performed to investigate the effects of socio-economic and demographic variables on the CCPI scores of countries. All explanatory variables were scaled. Backwards model selection using AIC was performed to obtain the simplest significant model. The model diagnostic plots were checked.

```{r AIC Selection Approach, echo=TRUE}
#Upper model
M1 <- lm(CCPI_Score ~ scale(School_Enrollment) + scale(Digital_Use) + 
              scale(School_Enrollment_Gender_Ratio) + 
              scale(Labour_Participation_Rate) + scale(Unemployment) + 
              scale(GDP_Growth) + scale(GDP_Capita)  + scale(Life_Expectancy) + 
              scale(Sex_Ratio) + scale(Population_Growth) + 
              scale(Population_Density), data = Final)
#Backwards model selection
M2<-step(M1, direction = "backward", scope = list(lower=~1, 
              upper=~scale(School_Enrollment) + scale(Digital_Use) + 
              scale(School_Enrollment_Gender_Ratio) + 
              scale(Labour_Participation_Rate) + scale(Unemployment) + 
              scale(GDP_Growth) + scale(GDP_Capita)  + scale(Life_Expectancy) + 
              scale(Sex_Ratio) + scale(Population_Growth) + 
              scale(Population_Density))) 

#Model Interpretation
summary(M2)

#Model diagnostic plots
plot(M2)
#Residual plots look good 
#QQ (normality plot) looks good

#Fitting a model without scaled variables (for plotting)
fit <- lm(CCPI_Score ~ School_Enrollment + Digital_Use + GDP_Capita + 
            Population_Growth, data = Final)

#Predicted values of CCPI score for School Enrollment (with other effects controlled)
G1 <- ggpredict(fit, terms = 'School_Enrollment')
#plotting using ggpredict (School Enrolment)
D1 <- ggplot(G1, aes(x, predicted)) + geom_line(colour = 'sienna1', size=1) + 
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.3) +
  ggtitle('CCPI Score vs Ratio of School Enrolment') + 
  xlab('Ratio of School Enrolment (%)') +
  ylab('CCPI Score') +  theme(plot.title = element_text(hjust=0.5, size = 12)) + 
  theme(axis.title = element_text(size=10)) + ylim(40,70) + xlim(100,110)

#Predicted values of CCPI score for GDP per Capita (with other effects controlled)
G2 <- ggpredict(fit, terms = 'GDP_Capita')
#plotting using ggpredict (GDP per Capita)
D2 <- ggplot(G2, aes(x, predicted)) + geom_line(colour = "sienna1", size=1) + 
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.3) +
  ggtitle('CCPI Score vs GDP per Capita') + xlab('GDP per Capita (US$)') +
  ylab('CCPI Score') + theme(plot.title = element_text(hjust=0.5, size = 12)) + 
  theme(axis.title = element_text(size=10)) + ylim(40,70) + xlim(30000,80000)

#Predicted values of CCPI score for digital usage (with other effects controlled)
G3 <- ggpredict(fit, terms = 'Digital_Use')
#plotting using ggpredict (GDP per Capita)
D3 <- ggplot(G3, aes(x, predicted)) + geom_line(colour = "lightseagreen", size=1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.3) +
  ggtitle('CCPI Score vs Digital Technology Usage') + 
  xlab('Digital Technology Usage (%)') +
  ylab('CCPI Score') + theme(plot.title = element_text(hjust=0.5, size = 12)) + 
  theme(axis.title = element_text(size=10)) + ylim(40,70) + xlim(60,90)

#Predicted values of CCPI score for population growth (with other effects controlled)
G4 <- ggpredict(fit, terms = 'Population_Growth')
#plotting using ggpredict (GDP per Capita)
D4 <- ggplot(G4, aes(x, predicted)) + 
  geom_line(colour = "lightseagreen", size=1) + 
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.3) +
  ggtitle('CCPI Score vs Population Growth') + 
  xlab('Population Growth (%)') +
  ylab('CCPI Score') +  theme(plot.title = element_text(hjust=0.5, size = 12)) + 
  theme(axis.title = element_text(size=10)) + ylim(40,70) + xlim(-0.25,1)

#combining the plots
plot_grid(D1, D2, D3, D4, labels = "AUTO", ncol = 2)
```
From the simplest significant regression model, School Enrollment, Digital Literacy, GDP per Capita and Population Growth were found to have a significant effect on CCPI Scores. All the model diagnostic plots satisfied the conditions of the linear model. Results further interpreted in the report. 
